Inference of Gravitational Lens Time Delays Using Gaussian Processes

### 2013-04-29 doc/Leon/ms.tex
Adding details

% Load some packages
\usepackage{graphicx} % This allows you to put figures in
\usepackage{natbib}   % This allows for relatively pain-free reference lists
\usepackage[left=3cm,top=3cm,right=3cm]{geometry} % The way I like the margins

@@

Suppose that the underlying QSO variability as a function of time is
described by:
\begin{eqnarray}
y_{\rm QSO}(t)
\end{eqnarray}
Then the light curves of the two images $A$ and $B$ will be given by:
\begin{eqnarray}
y_A(t) &=& y_{\rm QSO}(t) + \mu_A(t)\\
y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t)
\end{eqnarray}
where $\mu_A(t)$ and $\mu_B(t)$ are perturbations due to microlensing.

The data for all images are in magnitudes:
\begin{eqnarray}
\yy = \{y_1, y_2, ..., y_N\}
\end{eqnarray}

The probability distribution for the data $\yy$ given the parameters
$\theta$
is a multivariate normal distribution, where the mean vector $\mm$
and the covariance matrix $\CC$ depend on the parameters:
\begin{eqnarray}
p(\yy | \theta) &=& \frac{1}{\sqrt{(2\pi)^N\det{\CC}}}
\exp\left[
-\frac{1}{2}
\left(
\yy - \mm
\right)^T
\CC^{-1}
\left(
\yy - \mm
\right)
\right]
\end{eqnarray}
Numerically, the logarithm of the determinant and the
$\CC^{-1}\left(\yy - \mm\right)$ term are implemented using
Cholesky decompositions.

@@

>>>>

\usepackage{dsfont}
% Load some packages
\usepackage{graphicx} % This allows you to put figures in
\usepackage{natbib}   % This allows for relatively pain-free reference lists
\usepackage[left=3cm,top=3cm,right=3cm]{geometry} % The way I like the margins

@@

Throughout this section we consider the measured signal in magnitudes,
rather than fluxes. Suppose that the underlying QSO variability as a function
of time is described by the following function:
\begin{eqnarray}
y_{\rm QSO}(t)
\end{eqnarray}
Then the noise-free light curves of the two images
$A$ and $B$ will be given by:
\begin{eqnarray}
y_A(t) &=& y_{\rm QSO}(t) + \mu_A(t)\\
y_B(t) &=& y_{\rm QSO}(t - \tau) + \mu_B(t).
\end{eqnarray}
The variability of image $A$ is the underlying QSO variability
plus a signal $\mu_A(t)$ that is due to microlensing or perhaps
other effects. The variability of image $B$ is given by
a lagged version of the QSO variability, plus another
perturbation $\mu_B(t)$ caused my microlensing.

The data are assumed to be noisy measurements of $y_A(t)$ and
$y_B(t)$ at various times. We denote the entire dataset (both images together)
as a vector of magnitude measurements:
\begin{eqnarray}
\yy = \{y_1, y_2, ..., y_N\}.
\end{eqnarray}

If we make the simplifying assumption that our prior knowledge about
$\mu_A(t)$, $\mu_B(t)$, and
$y_{\rm QSO}(t)$ can be described by Gaussian Processes \citep{rasmussen},
then the probability distribution for the data $\yy$ given the parameters
$\theta$
is a multivariate normal distribution, where the mean vector $\mm$
and the covariance matrix $\CC$ depend on the parameters:
\begin{eqnarray}
p(\yy | \theta) &=& \frac{1}{\sqrt{(2\pi)^N\det{\CC}}}
\exp\left[
-\frac{1}{2}
\left(
\yy - \mm
\right)^T
\CC^{-1}
\left(
\yy - \mm
\right)
\right]
\end{eqnarray}
The elements of the mean vector are just the mean magnitudes of the
appropriate images:
\begin{eqnarray}
m_i = 
\end{eqnarray}

The covariance matrix elements are given by:
\begin{eqnarray}
C_{ij} &=& \textnormal{Cov}(y_i, y_j) \\
&=& \mathds{E}\left[
\left(y_i - m_i\right)
\left(y_j - m_j\right)
\right]
\end{eqnarray}

Numerically, the logarithm of the determinant and the
$\CC^{-1}\left(\yy - \mm\right)$ term are implemented using
Cholesky decompositions.

@@

\begin{thebibliography}{99} 
\bibitem[\protect\citeauthoryear{Rasmussen \& Williams}{2006}]{rasmussen} Rasmussen C., Williams C., Gaussian Processes for Machine Learning, The MIT Press, 2006
\end{thebibliography}
